---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse) #bringing in tidyverse
```

calling the data data
```{r}
data <- read_csv("Data/2021_LoL_esports_match_data_from_OraclesElixir_20211004.csv")
```

A definitions page is available here: https://oracleselixir.com/definitions
All relevant, included tournaments are here: https://oracleselixir.com/stats/players/byTournament, EM = EUM
Defining additional terms like Major Leagues being the LCS, LCK, LEC, LPL and the international tournament MSI(Mid-Season Invitational)
```{r}
Majors <- c("LCS","LEC","LCK","LPL","MSI")
```
Minors will be written as not in Majors


```{r}
data %>% group_by(gameid) %>% count()
```
Grouping the data to verify what they are, 12 on each is correct (1 for each player and 1 for team 2 teams + 10 players=12 instances)

```{r}
data %>% group_by(gameid) %>% count() %>% arrange(desc(n()))
```
Verified by flipping the arrange that all are 12 (and no NAs for game count)

```{r}
data %>% group_by(datacompleteness) %>% count()
```
Checking what values appear under datacompleteness

```{r}
NACompletes <- data %>% filter(is.na(datacompleteness))
```
Just wanted to see what some of the NAs were, some of this might still be usable, but I might just focus on what follows

This is likely what I will focus on using
```{r}
CompleteData <- data %>% filter(datacompleteness == "complete")
```
This represents 93,648/133,428 rows or about 61% of the data is complete, another 9% is partial (not shown)

Filtering for games to find the max number of turret plates, removed desc to check for lowest
```{r}
data %>% group_by(gameid) %>% filter(!is.na(turretplates)) %>% select(gameid, datacompleteness, turretplates) %>% arrange(desc(turretplates))
```
0-15 is max as seen in game. Only outer turrets have plates, there are 3 outer turrets with 5 plates each so 
15 plates is a major early lead, only happened 3 times

Curious as to which regions those three games were.
```{r}
CompleteData %>% filter((playerid == "100" | playerid == "200") & (gameid == "ESPORTSTMNT03_2074070" | gameid == "ESPORTSTMNT06_1881146" | gameid == "ESPORTSTMNT02_2151778")) %>%
  select(gameid, league, date, turretplates)
```
Another way to do this would be to filter not for playerid being a team but for turretplates equaling 15
The three leagues where absolute dominance occurred were HM (Hitpoint Masters) EM (EU Masters) and LCSA (League Championship Series Academy), in other words all "minor league" games

More Questions:

*Which player has the highest winrate (they may change teams)?
```{r}
CompleteData %>% group_by(player, league) %>% summarize(WR = sum(result)/n(), sum(result), totalG = n()) %>% arrange(desc(WR), desc(totalG))

CompleteData %>% filter(player == "113") #checking math manually, so not much code here
```

Chim, Competition9, and Jozy all had a 100% Winrate but they were in amateur leagues. 16 wins could be a clean sweep of a tournament, it could be interesting to see where their teammates are. But more importantly, Bo is the only major league player (sans the player Xiaobai who was subbed in for a single game at an international tournament, it was likely his team that carried him) with a 100% Winrate and he has 10 games, which is pretty impressive, especially for a powerhouse region like the LPL. 

```{r}
CompleteData %>% filter(is.na(player)) %>% group_by(league) %>% summarize(totalG = n()/2) %>% arrange(desc(totalG))
#I am immediately faced with a problem here, I have to filter where player is not na because the sets of data of just teams is NA so every team in each league balances each other out to 50% win rates, but I am keeping this here because it is a good way I think to get totals of games in each league. I am trimming it though.
```
NOTE ABOUT THE ABOVE TABLE: TOTAL GAMES IS HALVED BECAUSE EACH MATCH HAS TWO TEAMS AND IT WAS COUNTING BOTH TEAMS IN A SINGLE GAME.


Time for what I was trying to tackle in the above table initially:
```{r}
CompleteData %>% filter(!is.na(player) & player != "unknown player") %>% group_by(player, league) %>% summarize(WR = sum(result)/n(), Wins = sum(result), totalG = n()) %>% arrange(desc(Wins), desc(WR))
```
Note: Found a new thing: even in complete data, "unknown player" is still a thing, redid it to remove that. 
This gives me a good idea of players who have won a lot of their games, but the winrates still rise and fall between them, so I am going to do two things:
*Highest winrate with 40+ games
*Highest winrate in major leagues (LCS, LCK, LEC, LPL, MSI (even though it is an event, it brings in international teams))
```{r}
CompleteData %>% filter(!is.na(player) & player != "unknown player") %>% group_by(player, league) %>% summarize(WR = sum(result)/n(), Wins = sum(result), totalG = n()) %>% filter(totalG > 40) %>% arrange(desc(Wins), desc(WR))

CompleteData %>% filter(!is.na(player) & player != "unknown player" & (league %in% Majors)) %>% group_by(player, league) %>% summarize(WR = sum(result)/n(), Wins = sum(result), totalG = n()) %>% filter(totalG > 30) %>% arrange(desc(Wins), desc(WR))
```



*Which team has the highest winrate (members may change, this evaluates the coach)?

*What things are most consistent in a fast win (A fast win will be defined as faster than 25 mins, I will need to create a new column for this)?

*DWG (Damwon Gaming, may also be referred to as DWGKIA) were world champions this year, how did they look during the year leading up to their win?

*This might be the hardest question, any teams in major leagues (LCS, LCK, LPL, LEC) perform on average worse than major academy leagues (LCSA, EM (EUM), LDL, LAS (LCK Academey))

My methodology was included in the final pitch and is a lot to repeat, but in short I will be focusing on Complete Data (and may go into partial)
I will be focusing on what aspects in the early game (turret plates, first blood, first dragon) have the surest effect on wins
and on the flip side, the games where an "upset" happened, which players were involved and what champions were involved?

###Tutorials###

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

